---
title: "AD Fractions"
author:
  - name: "Emir, Pal, and Steph"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-CBD-exosome-proteins.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "AD-fractions.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-CBD-exosome-proteins](https://github.com/eturkes/AD-CBD-exosome-proteins).*
*Please email for access.*

The data here will be referenced using the name `fractions`.
It is a label-free mass spec dataset consisting of 1 Alzheimer's Disease (AD) human brain tissue donor where extracellular vesicles (EVs) were isolated and fractionated into 8 fractions.
Each sample in the dataset represents a fraction from this single donor.

```{r}
# Some standard boilerplate.
# --------------------------
# Copyright 2019-2021 Emir Turkes, Guar Pallavi, Stephanie Fowler, UK DRI at UCL, Columbia
# University Medical Center
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
packages <- c(
  "conflicted", "circlize", "dplyr", "DEP", "SummarizedExperiment", "scales", "ggplot2", "DT",
  "ComplexHeatmap", "EnhancedVolcano", "GSEABase", "GSVA", "RColorBrewer", "limma", "tibble"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("utils.R"))
conflict_prefer("rowMedians", "Biobase", quiet = TRUE)
conflict_prefer("which", "BiocGenerics", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)

data_name <- "fractions"
`%notin%` <- Negate(`%in%`)
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up data.

cache_dir <- file.path("..", "tmp", "cache", data_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ----------------------------------------------------------
# --------------------------
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
