---
title: "Dataset 2 AD CTRL"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-CBD-exosome-proteins.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "dataset2-AD-CTRL.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-CBD-exosome-proteins](https://github.com/eturkes/AD-CBD-exosome-proteins). Please email me for access.*

The data here is derived from @dataset2 and will be referenced using the name `dataset2`.

```{r}
# Copyright 2019-2020 Emir Turkes
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
options(stringsAsFactors = FALSE)
packages <- c(
  "RColorBrewer", "dplyr", "DEP", "SummarizedExperiment", "ggplot2",
  "DT", "GSEABase", "GSVA", "limma"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "utils.R"))

data_name <- "dataset2"
color <- colorRampPalette(rev(brewer.pal(7, "RdBu")))(100)
color_legend <- c("lightgray", "darkgray")
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path(getwd(), "..", "assets") # Backed up data.

cache_dir <- file.path(getwd(), "..", "tmp", "cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
```

# All Data

## Prep

```{r}
data <- read.csv(file.path(assets_dir, data_name, "proteinGroups.txt"), sep = "\t")

# Reverse order of sample columns to make DE analysis more intuitive later.
col_start <- grep("AD_5333", colnames(data))
for (i in 1:length(col_start)) {
  section1_end <- col_start[i] - 1
  col_end <- col_start[i] + 8
  section2_start <- col_end + 1
  data <- data.frame(
    data[ , 1:section1_end], rev(data[ , col_start[i]:col_end]), data[ , section2_start:ncol(data)]
  )
}

data <- filter(data, Reverse != "+", Potential.contaminant != "+")
data <- make_unique(data, "Gene.names", "Protein.IDs", ";")
LFQ_columns <- grep("Intensity.", colnames(data))
label = gsub("Peptides.", "", colnames(data)[13:21])
experimental_design <- data.frame(
  label = label, condition = c(rep("Contr", 4), rep("AD", 5)), replicate = label
)

data <- make_se(data, LFQ_columns, experimental_design)
# colnames(data) <- gsub("^.*?_", "", colnames(data)) # Remove redundant labels. # TODO: Fix this.
data
```

**Unaltered data, already log2 transformed.**

```{r}
datatable_download(assay(data))
```

**After removal of log transformation.**

```{r}
datatable_download(2 ^ assay(data))
```

## QC

### Histogram

```{r}
hist(assay(data), n = 100)
```

### Missing Value Removal

Retained proteins are those identified in 2 out of 3 replicates of at least one condition, a standard procedure for less stringent filtering.

```{r}
plot_numbers(data)
plot_frequency(data)
plot_missval(data)
data <- filter_proteins(data, "condition", 1)
plot_missval(data)
plot_numbers(data)
plot_frequency(data)
```

**Data before normalization but after removal of missing values.**

```{r}
datatable_download(assay(data))
```

### Normalization

Normalize data using scaling factors based on known concentrations of total protein in experiment, then log2 transform.

```{r}
concentrations <- c(1.61, 1.99, 1.92, 2.43, 1.07, 1.26, 1.43, NA, 1.41)
multipliers <- min(concentrations, na.rm = TRUE) / concentrations
missing_multiplier <- median(multipliers[5:length(multipliers)], TRUE)
multipliers <- c(multipliers[1:7], missing_multiplier, multipliers[9])

meanSdPlot(data)
data_norm <- data
assay(data_norm) <- 2 ^ assay(data_norm) # Undo log transformation.
for (i in seq_along(multipliers)) {
  assay(data_norm)[ , i] <- assay(data_norm)[ , i] * multipliers[i]
}
assay(data_norm) <- log2(assay(data_norm))
meanSdPlot(data_norm)
plot_normalization(data, data_norm)
data <- data_norm
```

**Data after normalization and log2 transformation.**

```{r}
datatable_download(assay(data))
```

### Missing Value Imputation

Using the QRILC method, which targets missing-not-at-random values AKA those that are likely missing due to being below the mass-spec quantification threshold.

```{r}
plot_detect(data)
set.seed(1)
data_imp <- impute(data, "QRILC")
plot_imputation(data, data_imp)
data <- data_imp
rm(data_imp)
```

**Data after missing value imputation.**
**This is final form used for differential expression analysis**

```{r}
datatable_download(assay(data))
```

### PCA

```{r}
plot_pca(data, n = round(dim(data)[1] * 0.1)) + scale_shape_manual(values = seq(0, length(label)))
```

## Differential Expression

```{r}
data <- add_rejections(test_diff(data, "control", "Contr"), lfc = 0)
data_results <- get_results(data)
nsig <- data_results %>% filter(significant) %>% nrow()
print(paste0(nsig, " significant proteins."))

plot_heatmap(data, "centered", TRUE, round((nsig -1) / 2), row_font_size = 9)
plot_cor(data, pal = "RdBu", pal_rev = TRUE)
plot_volcano(data, "AD_vs_Contr", 4, adjusted = TRUE)

plot_single(data, c("MAPT", "APOE"), "centered") + xlab(element_blank())

datatable_download(data_results)
```

## Gene Set Enrichment

We perform gene set enrichment using GSVA [@hanzelmannGSVAGeneSet2013].

### Cell-types

We use a custom GMT file compiled from [PanglaoDB](https://panglaodb.se/).
The results below are **NOT** filtered for significance.

```{r}
gene_set_name <- "panglao-brain-ctypes"
gene_sets <- getGmt(file.path(assets_dir, "gene-sets", paste0(gene_set_name, ".gmt")))
gene_sets <- filterGeneSets(gene_sets, 5, 500)
gene_sets

gsva <- gsva(assay(data), gene_sets)
datatable_download(gsva)

Contr <- data[ , grep("Contr", colnames(data))]
AD <- data[ , grep("AD", colnames(data))]
conditions <- c("Contr", "AD")
comb_mat <- cbind(assay(Contr), assay(AD))
rownames(comb_mat) <- rownames(data)
groups <- factor(rep(rev(conditions), c(dim(Contr)[2], dim(AD)[2])))
levels(groups) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]
colnames <- gsub("^.*?_", "", colnames(gsva))

gene_set_clustering <- hclust(as.dist(1 - cor(t(gsva), method = "pearson")), "complete")
heatmap(
  gsva,
  ColSideColors = sample_color_map,
  margins = c(6, 10),
  col = color,
  ylab = gene_set_name,
  labCol = colnames,
  labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(gsva))),
  Rowv = as.dendrogram(gene_set_clustering),
  Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

### GO Biological Processes

We use a GO Biological Process (BP) GMT file from [MSigDB](https://www.gsea-msigdb.org/gsea/msigdb/index.jsp).
The results below are filtered for significance.

```{r}
gene_set_name <- "broad-GO-BP"
gene_sets <- getGmt(file.path(assets_dir, "gene-sets", paste0(gene_set_name, ".gmt")))
gene_sets <- filterGeneSets(gene_sets, 5, 500)
gene_sets

gsva <- gsva(assay(data), gene_sets)
datatable_download(gsva)

Contr <- data[ , grep("Contr", colnames(data))]
AD <- data[ , grep("AD", colnames(data))]
conditions <- c("Contr", "AD")
comb_mat <- cbind(assay(Contr), assay(AD))
rownames(comb_mat) <- rownames(data)
groups <- factor(rep(rev(conditions), c(dim(Contr)[2], dim(AD)[2])))
levels(groups) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]
colnames <- gsub("^.*?_", "", colnames(gsva))

fit <- eBayes(lmFit(gsva, model.matrix(~ groups)))
DE_gene_sets <- topTable(fit, paste0("groups", conditions[2]), Inf, p.value = 0.05)
datatable_download_exp(DE_gene_sets)

gsva <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva)
gene_set_clustering <- hclust(as.dist(1 - cor(t(gsva), method = "pearson")), "complete")
heatmap(
  gsva,
  ColSideColors = sample_color_map,
  margins = c(6, 25),
  col = color,
  labCol = colnames,
  labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(gsva))),
  Rowv = as.dendrogram(gene_set_clustering),
  Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")

print("RESPONSE_TO_MISFOLDED_PROTEIN")
gene_sets[["GO_RESPONSE_TO_MISFOLDED_PROTEIN"]]@shortDescription
gene_sets[["GO_RESPONSE_TO_MISFOLDED_PROTEIN"]]@geneIds

genes <- data[rownames(data) %in% gene_sets[["GO_RESPONSE_TO_MISFOLDED_PROTEIN"]]@geneIds, ]
gene_set_clustering <- hclust(as.dist(1 - cor(t(assay(genes)), method = "pearson")), "complete")
heatmap(
  assay(genes),
  ColSideColors = sample_color_map,
  margins = c(6, 10),
  col = color,
  labCol = colnames,
  labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(genes))),
  Rowv = as.dendrogram(gene_set_clustering),
  Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")

plot_single(data, c("VCP", "CLU"), "centered") + xlab(element_blank())
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
