---
title: "Imputation Simulation Comparison"
author:
  - name: "Emir Turkes and Stephanie Fowler"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-exosome-characterisation.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "imputation-simulation-comparison.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
.tocify-subheader .tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 45px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-exosome-characterisation](https://github.com/eturkes/AD-exosome-characterisation).*
*Please email for access.*

```{r}
# Some standard boilerplate.
# --------------------------
#    This file is part of AD-exosome-characterisation.
#    Copyright (C) 2022  Emir Turkes, Stephanie Fowler, UK DRI at UCL, Columbia
#    University Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# --------------------------------------------
packages <- c(
  "conflicted", "SummarizedExperiment", "limma", "khroma", "stringr", "DT", "volcano3D",
  "VennDiagram", "RColorBrewer", "htmltools", "ComplexHeatmap", "circlize"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("utils.R"))

`%notin%` <- Negate(`%in%`)

data_name <- "oxford-ms-data-april-2022"
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up data.

cache_dir <- file.path("..", "tmp", "cache", data_name, "imputation-simulation-comparison")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# ----------------------------------------------------------
# --------------------------
```

# Prep

```{r}
data_list <- list(
  "imputed" = readRDS(file.path(cache_dir, "..", "imputation-simulation", "imputed.rds")),
  "unimputed" = readRDS(file.path(cache_dir, "..", "imputation-simulation", "unimputed.rds")),
  "complete" = readRDS(file.path(cache_dir, "..", "imputation-simulation", "complete.rds"))
)

limma_design <- model.matrix(~ 0 + data_list[[1]]$groups)
colnames(limma_design) <- unique(data_list[[1]]$groups)
contrast_mat <- makeContrasts(F1_3-F4_6, F1_3-F7_8, F4_6-F7_8, levels = limma_design)

colour <- colour("vibrant")(7)
colour <- append("#DDAA33", colour)
colour <- colour[c(1, 2, 5, 3, 4, 7, 6, 8)]
names(colour) <- unique(sub("\\.", " ", data_list[[1]]$condition))

intensity_range <- range(
  rbind(rbind(assay(data_list[[1]]), assay(data_list[[2]]))), assay(data_list[[3]]), na.rm = TRUE
)
heatmap_colour <- colorRamp2(
  c(
    intensity_range[1], intensity_range[1] + (intensity_range[2] - intensity_range[1]) / 2,
    intensity_range[2]
  ),
  c("#0077BB", "#FFFFFF", "#CC3311")
)
```

# DEP Pipeline

## `r str_to_title(names(data_list)[1])`

```{r}
idx <- 1

fit <- lmFit(assay(data_list[[idx]]), limma_design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "global")
write.fit(
  cont_fit, tests, file.path(cache_dir, paste(names(data_list)[idx], "results.tsv", sep = "_")),
  adjust = "BH", method = "global", F.adjust = "BH"
)
results <- read.delim(file.path(cache_dir, paste(names(data_list)[idx], "results.tsv", sep = "_")))
rownames(results) <- results$X
results <- results[ , -1]
results <- results[order(results$F.p.value), ]

datatable_download_exp(results)
```

### F1_3

```{r}
tests_up <- tests[
  tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 0 |
    tests[ , 1] == 1 & tests[ , 2] == 0 & tests[ , 3] == 0 |
    tests[ , 1] == 0 & tests[ , 2] == 1 & tests[ , 3] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F1_3 <- rownames(up)[1]
F1_3_imputed <- rownames(up)
datatable_download_exp(up)
```

### F1_3 + F4_6

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 2] == 1 & tests[ , 3] == 1 |
    tests[ , 1] == 0 & tests[ , 2] == 1 & tests[ , 3] == 1 |
    tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F1_3_plus_F4_6 <- rownames(up)[1]
datatable_download_exp(up)
```

### F4_6

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 3] == 1 & tests[ , 2] == 0 |
    tests[ , 1] == -1 & tests[ , 3] == 0 & tests[ , 2] == 0 |
    tests[ , 1] == 0 & tests[ , 3] == 1 & tests[ , 2] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F4_6 <- rownames(up)[1]
F4_6_imputed <- rownames(up)
datatable_download_exp(up)
```

### F4_6 + F7_8

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 3] == -1 & tests[ , 2] == -1 |
    tests[ , 1] == -1 & tests[ , 3] == 0 & tests[ , 2] == -1 |
    tests[ , 1] == -1 & tests[ , 3] == 1 & tests[ , 2] == -1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F4_6_plus_F7_8 <- rownames(up)[1]
datatable_download_exp(up)
```

### F7_8

```{r}
tests_up <- tests[
  tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 1] == 0 |
    tests[ , 2] == -1 & tests[ , 3] == 0 & tests[ , 1] == 0 |
    tests[ , 2] == 0 & tests[ , 3] == -1 & tests[ , 1] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F7_8 <- rownames(up)[1]
F7_8_imputed <- rownames(up)
datatable_download_exp(up)
```

### F7_8 + F1_3

```{r}
tests_up <- tests[
  tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 1] == 1 |
    tests[ , 2] == 0 & tests[ , 3] == -1 & tests[ , 1] == 1 |
    tests[ , 2] == 1 & tests[ , 3] == -1 & tests[ , 1] == 1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F7_8_plus_F1_3 <- rownames(up)[1]
datatable_download_exp(up)
```

### Radial and 3D Volcano Plots

```{r, dev = "svglite"}
DEPs_main <- c(F1_3, F4_6, F7_8)
DEPs <- c(DEPs_main, F1_3_plus_F4_6, F4_6_plus_F7_8, F7_8_plus_F1_3)

mat <- assay(data_list[[idx]])
colnames(mat) <- data_list[[idx]]$nice_names
experimental_design <- data.frame(groups = data_list[[idx]]$groups)
rownames(experimental_design) <- data_list[[idx]]$nice_names

results <- results[match(rownames(mat), rownames(results)), ]

p_mat <- as.matrix(
  data.frame(
    F.p.value = results$F.p.value,
    P.value.F1_3...F4_6 = results$P.value.F1_3...F4_6,
    P.value.F1_3...F7_8 = results$P.value.F1_3...F7_8,
    P.value.F4_6...F7_8 = results$P.value.F4_6...F7_8
  )
)

p_adj_mat <- as.matrix(
  data.frame(
    F.p.value = results$F.p.value,
    P.value.adj.F1_3...F4_6 = results$P.value.adj.F1_3...F4_6,
    P.value.adj.F1_3...F7_8 = results$P.value.adj.F1_3...F7_8,
    P.value.adj.F4_6...F7_8 = results$P.value.adj.F4_6...F7_8
  )
)

polar_main <- polar_coords(
  experimental_design$groups,
  t(mat),
  p_mat,
  p_adj_mat,
  scheme = c(
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[1]],
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[3]],
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[4]],
    colour("vibrant")(7)[[7]]
  ),
  labs = unique(experimental_design$groups)
)

radial_ggplot(polar_main, 2, label_rows = DEPs_main, colour_code_labels = TRUE, marker_size = 2)
```

#### Interactive

```{r, dpi = 96}
radial_plotly(polar_main, 2)
```

```{r, dpi = 96}
volcano3D(polar_main, 2)
```

## `r str_to_title(names(data_list)[2])`

```{r}
idx <- 2

fit <- lmFit(assay(data_list[[idx]]), limma_design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "global")
write.fit(
  cont_fit, tests, file.path(cache_dir, paste(names(data_list)[idx], "results.tsv", sep = "_")),
  adjust = "BH", method = "global", F.adjust = "BH"
)
results <- read.delim(file.path(cache_dir, paste(names(data_list)[idx], "results.tsv", sep = "_")))
rownames(results) <- results$X
results <- results[ , -1]
results <- results[order(results$F.p.value), ]

datatable_download_exp(results)
```

### F1_3

```{r}
tests_up <- tests[
  tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 0 |
    tests[ , 1] == 1 & tests[ , 2] == 0 & tests[ , 3] == 0 |
    tests[ , 1] == 0 & tests[ , 2] == 1 & tests[ , 3] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F1_3 <- rownames(up)[1]
F1_3_unimputed <- rownames(up)
datatable_download_exp(up)
```

### F1_3 + F4_6

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 2] == 1 & tests[ , 3] == 1 |
    tests[ , 1] == 0 & tests[ , 2] == 1 & tests[ , 3] == 1 |
    tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F1_3_plus_F4_6 <- rownames(up)[1]
datatable_download_exp(up)
```

### F4_6

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 3] == 1 & tests[ , 2] == 0 |
    tests[ , 1] == -1 & tests[ , 3] == 0 & tests[ , 2] == 0 |
    tests[ , 1] == 0 & tests[ , 3] == 1 & tests[ , 2] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F4_6 <- rownames(up)[1]
F4_6_unimputed <- rownames(up)
datatable_download_exp(up)
```

### F4_6 + F7_8

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 3] == -1 & tests[ , 2] == -1 |
    tests[ , 1] == -1 & tests[ , 3] == 0 & tests[ , 2] == -1 |
    tests[ , 1] == -1 & tests[ , 3] == 1 & tests[ , 2] == -1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F4_6_plus_F7_8 <- rownames(up)[1]
datatable_download_exp(up)
```

### F7_8

```{r}
tests_up <- tests[
  tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 1] == 0 |
    tests[ , 2] == -1 & tests[ , 3] == 0 & tests[ , 1] == 0 |
    tests[ , 2] == 0 & tests[ , 3] == -1 & tests[ , 1] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F7_8 <- rownames(up)[1]
F7_8_unimputed <- rownames(up)
datatable_download_exp(up)
```

### F7_8 + F1_3

```{r}
tests_up <- tests[
  tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 1] == 1 |
    tests[ , 2] == 0 & tests[ , 3] == -1 & tests[ , 1] == 1 |
    tests[ , 2] == 1 & tests[ , 3] == -1 & tests[ , 1] == 1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F7_8_plus_F1_3 <- rownames(up)[1]
datatable_download_exp(up)
```

### Radial and 3D Volcano Plots

```{r, dev = "svglite"}
DEPs_main <- c(F1_3, F4_6, F7_8)
DEPs <- c(DEPs_main, F1_3_plus_F4_6, F4_6_plus_F7_8, F7_8_plus_F1_3)

mat <- assay(data_list[[idx]])
colnames(mat) <- data_list[[idx]]$nice_names
experimental_design <- data.frame(groups = data_list[[idx]]$groups)
rownames(experimental_design) <- data_list[[idx]]$nice_names

results <- results[match(rownames(mat), rownames(results)), ]

p_mat <- as.matrix(
  data.frame(
    F.p.value = results$F.p.value,
    P.value.F1_3...F4_6 = results$P.value.F1_3...F4_6,
    P.value.F1_3...F7_8 = results$P.value.F1_3...F7_8,
    P.value.F4_6...F7_8 = results$P.value.F4_6...F7_8
  )
)

p_adj_mat <- as.matrix(
  data.frame(
    F.p.value = results$F.p.value,
    P.value.adj.F1_3...F4_6 = results$P.value.adj.F1_3...F4_6,
    P.value.adj.F1_3...F7_8 = results$P.value.adj.F1_3...F7_8,
    P.value.adj.F4_6...F7_8 = results$P.value.adj.F4_6...F7_8
  )
)

polar_main <- polar_coords(
  experimental_design$groups,
  t(mat),
  p_mat,
  p_adj_mat,
  scheme = c(
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[1]],
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[3]],
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[4]],
    colour("vibrant")(7)[[7]]
  ),
  labs = unique(experimental_design$groups)
)

radial_ggplot(polar_main, 2, label_rows = DEPs_main, colour_code_labels = TRUE, marker_size = 2)
```

#### Interactive

```{r, dpi = 96}
radial_plotly(polar_main, 2)
```

```{r, dpi = 96}
volcano3D(polar_main, 2)
```

## `r str_to_title(names(data_list)[3])`

```{r}
idx <- 3

fit <- lmFit(assay(data_list[[idx]]), limma_design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "global")
write.fit(
  cont_fit, tests, file.path(cache_dir, paste(names(data_list)[idx], "results.tsv", sep = "_")),
  adjust = "BH", method = "global", F.adjust = "BH"
)
results <- read.delim(file.path(cache_dir, paste(names(data_list)[idx], "results.tsv", sep = "_")))
rownames(results) <- results$X
results <- results[ , -1]
results <- results[order(results$F.p.value), ]

datatable_download_exp(results)
```

### F1_3

```{r}
tests_up <- tests[
  tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 0 |
    tests[ , 1] == 1 & tests[ , 2] == 0 & tests[ , 3] == 0 |
    tests[ , 1] == 0 & tests[ , 2] == 1 & tests[ , 3] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F1_3 <- rownames(up)[1]
F1_3_complete <- rownames(up)
datatable_download_exp(up)
```

### F1_3 + F4_6

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 2] == 1 & tests[ , 3] == 1 |
    tests[ , 1] == 0 & tests[ , 2] == 1 & tests[ , 3] == 1 |
    tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F1_3_plus_F4_6 <- rownames(up)[1]
datatable_download_exp(up)
```

### F4_6

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 3] == 1 & tests[ , 2] == 0 |
    tests[ , 1] == -1 & tests[ , 3] == 0 & tests[ , 2] == 0 |
    tests[ , 1] == 0 & tests[ , 3] == 1 & tests[ , 2] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F4_6 <- rownames(up)[1]
F4_6_complete <- rownames(up)
datatable_download_exp(up)
```

### F4_6 + F7_8

```{r}
tests_up <- tests[
  tests[ , 1] == -1 & tests[ , 3] == -1 & tests[ , 2] == -1 |
    tests[ , 1] == -1 & tests[ , 3] == 0 & tests[ , 2] == -1 |
    tests[ , 1] == -1 & tests[ , 3] == 1 & tests[ , 2] == -1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F4_6_plus_F7_8 <- rownames(up)[1]
datatable_download_exp(up)
```

### F7_8

```{r}
tests_up <- tests[
  tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 1] == 0 |
    tests[ , 2] == -1 & tests[ , 3] == 0 & tests[ , 1] == 0 |
    tests[ , 2] == 0 & tests[ , 3] == -1 & tests[ , 1] == 0,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F7_8 <- rownames(up)[1]
F7_8_complete <- rownames(up)
datatable_download_exp(up)
```

### F7_8 + F1_3

```{r}
tests_up <- tests[
  tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 1] == 1 |
    tests[ , 2] == 0 & tests[ , 3] == -1 & tests[ , 1] == 1 |
    tests[ , 2] == 1 & tests[ , 3] == -1 & tests[ , 1] == 1,
]
up <- results[rownames(results) %in% rownames(tests_up), ]
F7_8_plus_F1_3 <- rownames(up)[1]
datatable_download_exp(up)
```

### Radial and 3D Volcano Plots

```{r, dev = "svglite"}
DEPs_main <- c(F1_3, F4_6, F7_8)
DEPs <- c(DEPs_main, F1_3_plus_F4_6, F4_6_plus_F7_8, F7_8_plus_F1_3)

mat <- assay(data_list[[idx]])
colnames(mat) <- data_list[[idx]]$nice_names
experimental_design <- data.frame(groups = data_list[[idx]]$groups)
rownames(experimental_design) <- data_list[[idx]]$nice_names

results <- results[match(rownames(mat), rownames(results)), ]

p_mat <- as.matrix(
  data.frame(
    F.p.value = results$F.p.value,
    P.value.F1_3...F4_6 = results$P.value.F1_3...F4_6,
    P.value.F1_3...F7_8 = results$P.value.F1_3...F7_8,
    P.value.F4_6...F7_8 = results$P.value.F4_6...F7_8
  )
)

p_adj_mat <- as.matrix(
  data.frame(
    F.p.value = results$F.p.value,
    P.value.adj.F1_3...F4_6 = results$P.value.adj.F1_3...F4_6,
    P.value.adj.F1_3...F7_8 = results$P.value.adj.F1_3...F7_8,
    P.value.adj.F4_6...F7_8 = results$P.value.adj.F4_6...F7_8
  )
)

polar_main <- polar_coords(
  experimental_design$groups,
  t(mat),
  p_mat,
  p_adj_mat,
  scheme = c(
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[1]],
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[3]],
    colour("vibrant")(7)[[7]], colour("vibrant")(7)[[4]],
    colour("vibrant")(7)[[7]]
  ),
  labs = unique(experimental_design$groups)
)

radial_ggplot(polar_main, 2, label_rows = DEPs_main, colour_code_labels = TRUE, marker_size = 2)
```

#### Interactive

```{r, dpi = 96}
radial_plotly(polar_main, 2)
```

```{r, dpi = 96}
volcano3D(polar_main, 2)
```

# DEP Comparison

```{r}
results_list <- list(
  "imputed" = read.delim(
    file.path(cache_dir, paste(names(data_list)[1], "results.tsv", sep = "_"))
  ),
  "unimputed" = read.delim(
    file.path(cache_dir, paste(names(data_list)[2], "results.tsv", sep = "_"))
  ),
  "complete" = read.delim(
    file.path(cache_dir, paste(names(data_list)[3], "results.tsv", sep = "_"))
  )
)
```

## Venn Diagram (All DEPs)

```{r}
venn_list <- list(
  "imputed" = results_list[[1]]$X[
    results_list[[1]]$X %in% c(F1_3_imputed, F4_6_imputed, F7_8_imputed)
  ],
  "unimputed" = results_list[[2]]$X[
    results_list[[2]]$X %in% c(F1_3_unimputed, F4_6_unimputed, F7_8_unimputed)
  ],
  "complete" = results_list[[3]]$X[
    results_list[[3]]$X %in% c(F1_3_complete, F4_6_complete, F7_8_complete)
  ]
)
venn_list <- lapply(venn_list, function(x) x[!is.na(x)])

plot <- venn.diagram(
  venn_list, NULL, TRUE,
  category.names = str_to_title(names(venn_list)), fill = brewer.pal(3, "Pastel2")
)
grid.draw(plot)

overlap <- calculate.overlap(venn_list)
```

## Heatmaps (All DEPs)

```{r, results = "asis"}
for (name in names(venn_list)) {
  cat("\n")
  cat("###", str_to_title(name), "\n")

  set.seed(1)
  mat <- assay(
    data_list[[name]][
      rownames(data_list[[name]]) %in% sample(venn_list[[name]], length(venn_list$complete)),
    ]
  )
  mat <- mat[order(as.numeric(gsub("\\D+", "", rownames(mat)))), ]
  colnames(mat) <- data_list[[name]]$nice_names
  set.seed(1)
  draw(
    Heatmap(
      mat,
      heatmap_colour,
      cluster_columns = FALSE,
      cluster_rows = FALSE,
      heatmap_legend_param = list(
        title = "log2 Intensity", direction = "horizontal",
        legend_width = unit(7.5, "cm")
      ),
      row_names_gp = gpar(fontsize = 10),
      column_names_gp = gpar(fontsize = 10),
      rect_gp = gpar(col = "grey", lwd = 1)
    ),
    heatmap_legend_side = "top"
  )

  cat('\n')
}
```

## Venn Diagram (True DEPs)

```{r}
venn_list_true <- list(
  "imputed" = results_list[[1]]$X[
    results_list[[1]]$X %in% c(F1_3_imputed, F4_6_imputed, F7_8_imputed)][
      as.numeric(
        gsub(
          "\\D+", "",
          results_list[[1]]$X[
            results_list[[1]]$X %in% c(F1_3_imputed, F4_6_imputed, F7_8_imputed)
          ]
        )
      ) <= 2250
    ],
  "unimputed" = results_list[[2]]$X[
    results_list[[2]]$X %in% c(F1_3_unimputed, F4_6_unimputed, F7_8_unimputed)][
      as.numeric(
        gsub(
          "\\D+", "",
          results_list[[2]]$X[
            results_list[[2]]$X %in% c(F1_3_unimputed, F4_6_unimputed, F7_8_unimputed)
          ]
        )
      ) <= 2250
    ],
  "complete" = results_list[[3]]$X[
    results_list[[3]]$X %in% c(F1_3_complete, F4_6_complete, F7_8_complete)][
      as.numeric(
        gsub(
          "\\D+", "",
          results_list[[3]]$X[
            results_list[[3]]$X %in% c(F1_3_complete, F4_6_complete, F7_8_complete)
          ]
        )
      ) <= 2250
    ]
)
venn_list_true <- lapply(venn_list_true, function(x) x[!is.na(x)])

plot <- venn.diagram(
  venn_list_true, NULL, TRUE,
  category.names = str_to_title(names(venn_list_true)), fill = brewer.pal(3, "Pastel2")
)
grid.draw(plot)

overlap <- calculate.overlap(venn_list_true)
```

## Heatmaps (True DEPs)

```{r, results = "asis"}
for (name in names(venn_list_true)) {
  cat("\n")
  cat("###", str_to_title(name), "\n")

  if (name == "complete") {
    set.seed(1)
    mat <- assay(
      data_list[[name]][
        rownames(data_list[[name]]) %in%
          sample(venn_list_true[[name]], length(venn_list_true$complete)),
      ]
    )
    mat <- mat[order(as.numeric(gsub("\\D+", "", rownames(mat)))), ]
  } else {
    keep <- intersect(venn_list_true$imputed, venn_list_true$unimputed)[
      intersect(venn_list_true$imputed, venn_list_true$unimputed) %notin% venn_list_true$complete
    ]
    set.seed(1)
    mat <- assay(
      data_list[[name]][
        rownames(data_list[[name]]) %in% sample(keep, length(venn_list_true$complete)),
      ]
    )
    mat <- mat[order(as.numeric(gsub("\\D+", "", rownames(mat)))), ]
  }
  colnames(mat) <- data_list[[name]]$nice_names
  set.seed(1)
  draw(
    Heatmap(
      mat,
      heatmap_colour,
      cluster_columns = FALSE,
      cluster_rows = FALSE,
      heatmap_legend_param = list(
        title = "log2 Intensity", direction = "horizontal",
        legend_width = unit(7.5, "cm")
      ),
      row_names_gp = gpar(fontsize = 10),
      column_names_gp = gpar(fontsize = 10),
      rect_gp = gpar(col = "grey", lwd = 1)
    ),
    heatmap_legend_side = "top"
  )

  cat('\n')
}
```

## Percent True DEP Captured

```{r}
df <- data.frame(
  Imputed = round(length(venn_list_true$imputed) / 2250 * 100, 2),
  Unimputed = round(length(venn_list_true$unimputed) / 2250 * 100, 2),
  Complete = round(length(venn_list_true$complete) / 2250 * 100, 2)
)
datatable_download(df)
```

## Accuracy

```{r}
df <- data.frame(
  Imputed = round(length(venn_list_true$imputed) / length(venn_list$imputed) * 100, 2),
  Unimputed = round(length(venn_list_true$unimputed) / length(venn_list$unimputed) * 100, 2),
  Complete = round(length(venn_list_true$complete) / length(venn_list$complete) * 100, 2)
)
datatable_download(df)
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
