---
title: "TAU Peptides"
author:
  - name: "Emir, Pal, and Steph"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-CBD-exosome-proteins.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "TAU-peptides.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-CBD-exosome-proteins](https://github.com/eturkes/AD-CBD-exosome-proteins).*
*Please email for access.*

The data here will be referenced using the name `fractions`.
It is a label-free mass spec dataset consisting of 1 Alzheimer's Disease (AD) human brain tissue donor where extracellular vesicles (EVs) were isolated and fractionated into 8 fractions.
Each sample in the dataset represents a fraction from this single donor.
In this analysis we look at individual peptide sequences, focusing on TAU protein.

```{r}
# Some standard boilerplate.
# --------------------------
# Copyright 2019-2021 Emir Turkes, Guar Pallavi, Stephanie Fowler, UK DRI at UCL, Columbia
# University Medical Center
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
packages <- c(
  "conflicted", "stringr", "DEP"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("utils.R"))

data_name <- "fractions"
`%notin%` <- Negate(`%in%`)
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up data.
results_dir <- file.path("..", "results")

cache_dir <- file.path("..", "tmp", "cache", data_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# ----------------------------------------------------------
# --------------------------
```

# Prep

Read in the data and organize into a `SummarizedExperiment` object.

```{r}
data <- read.delim(file.path(assets_dir, data_name, "peptides.txt"))

# Use information from proteinGroups.txt to remove unwanted peptides.
# -------------------------------------------------------------------
cat("Number of peptides before removal of contaminants and decoys:", nrow(data))
protein_groups <- read.delim(file.path(assets_dir, data_name, "proteinGroups.txt"))
protein_groups <- protein_groups[
  protein_groups$Reverse == "+" | protein_groups$Potential.contaminant == "+",
]
remove <- unique(unlist(str_split(protein_groups$Peptide.IDs, ";")))
data <- data[data$id %notin% remove, ]
cat("Number of peptides after removal of contaminants and decoys:", nrow(data))
# -------------------------------------------------------------------

data$name <- paste(data$Gene, data$Sequence, sep = "_")
data$ID <- data$Proteins
LFQ_columns <- grep("LFQ.intensity.", colnames(data))
label <- gsub("LFQ.intensity.", "", colnames(data)[LFQ_columns])
experimental_design <- data.frame(
  label = label, condition = c(rep("fraction", 8)), replicate = label
)

data <- make_se(data, LFQ_columns, experimental_design)
data
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
