---
title: "Zetaview Analysis Concentration"
author:
  - name: "Emir Turkes and Stephanie Fowler"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-exosome-characterisation.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "zetaview-analysis-concentration.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-exosome-characterisation](https://github.com/eturkes/AD-exosome-characterisation).*
*Please email for access.*

The data here will be referenced using the name `zetaview`.
It is particle count data from ZetaView of exosomes derived from post-mortem brains of 8 donors with Alzheimer's Disease (AD).
Exosomes were fractionated into 8 fractions (plus 10k pellet, which are excluded from downstream analysis) for each donor.
Three technical replicates per fraction was run through the Zetaview.

```{r}
# Some standard boilerplate.
# --------------------------
#    This file is part of AD-exosome-characterisation.
#    Copyright (C) 2022  Emir Turkes, Stephanie Fowler, UK DRI at UCL, Columbia
#    University Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# --------------------------------------------
packages <- c("conflicted", "DT", "ggplot2", "reshape2", "scales", "ggbeeswarm", "khroma")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("utils.R"))

data_name <- "zetaview"
`%notin%` <- Negate(`%in%`)

colour <- colour("vibrant")(7)
colour <- append("#DDAA33", colour)
colour <- colour[c(1, 2, 5, 3, 4, 7, 6, 8)]
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up data.

cache_dir <- file.path("..", "tmp", "cache", data_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300, dev = "svglite")
# ----------------------------------------------------------
# --------------------------
```

# Prep

The data is spread out across many files, so we first aggregate them to obtain a single data object with counts per particle size.
We also create variations of this object that are aggregated by the mean of values by donor and fraction as well as various normalisation schemes found in following sections.

```{r}
file_list <- list.files(file.path(assets_dir, data_name, "raw-data"))
file_list <- file_list[-grep("10K", file_list)] # Remove 10K pellets.

data <- data.frame(matrix(nrow = 1200, ncol = length(file_list)) + 1)

for (i in seq_along(file_list)) {
  file <- read.delim(
    file.path(assets_dir, data_name, "raw-data", file_list[i]), skip = 70
  )
  dilution <- read.delim(
    file.path(assets_dir, data_name, "raw-data", file_list[i]), FALSE, nrows = 1, skip = 24
  )
  dilution <- dilution$V7
  if (i == 1) {
    data[ , i] <- file[1:1200, 1]
  }
  data[ , i + 1] <- file[1:1200, 3] * dilution
}

colnames(data) <- c(
  "size",
    paste(
      sub("^[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sub("^[^_]*_[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sub("^[^_]*_[^_]*_[^_]*_[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sep = "_"
  )
)

non_zero <- which(rowSums(data[ , -1]) > 0)
data <- data[min(non_zero):max(non_zero), ]

# Scaling factor based on tissue amount used.
# -------------------------------------------
scaling_factor <- data.frame(amount = c(1.25, 1.29, 1.21, 1.19, 1.05, 1.01, 1.15, 1.2))
scaling_factor <- 1 / scaling_factor
scaling_factor <- scaling_factor / 2e13
rownames(scaling_factor) <- unique(sub("^([^_]*).*", "\\1", colnames(data[ , -1])))
# -------------------------------------------

data_scaled <- data
for (i in seq_along(rownames(scaling_factor))) {
  for (j in seq_along(colnames(data_scaled))) {
    if (sub("^([^_]*).*", "\\1", colnames(data))[j] == rownames(scaling_factor)[i])
      data_scaled[j] <- data_scaled[j] * scaling_factor[i, ]
  }
}

data_norm <- data
for (i in seq_along(colnames(data_norm))) {
  if (i != 1) {
    data_norm[i] <- (data_norm[i] / sum(data_norm[i])) * 100
  }
}

data_mode <- data
for (i in seq_along(colnames(data_mode))) {
  if (i != 1) {
    data_mode[i] <- (data_mode[i] / max(data_mode[i]))
  }
}

names <- unique(sub("_[^_]+$", "", colnames(data[ , -1])))
donor_fraction_data <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
donor_fraction_data$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("_[^_]+$", "", colnames(data)) %in% names[i])]
  donor_fraction_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data) <- c("size", names)

names <- unique(sub("_[^_]+$", "", colnames(data_scaled[ , -1])))
donor_fraction_data_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
donor_fraction_data_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[ , which(sub("_[^_]+$", "", colnames(data_scaled)) %in% names[i])]
  donor_fraction_data_scaled[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data_scaled) <- c("size", names)

names <- unique(sub("_[^_]+$", "", colnames(data_norm[ , -1])))
donor_fraction_data_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
donor_fraction_data_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[ , which(sub("_[^_]+$", "", colnames(data_norm)) %in% names[i])]
  donor_fraction_data_norm[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data_norm) <- c("size", names)

names <- unique(sub("_[^_]+$", "", colnames(data_mode[ , -1])))
donor_fraction_data_mode <- data.frame(matrix(nrow = nrow(data_mode), ncol = length(names) + 1))
donor_fraction_data_mode$X1 <- data_mode$size
for (i in seq_along(names)) {
  data_sub <- data_mode[ , which(sub("_[^_]+$", "", colnames(data_mode)) %in% names[i])]
  donor_fraction_data_mode[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data_mode) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data[ , -1])))
donor_data <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
donor_data$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("^([^_]*).*", "\\1", colnames(data)) %in% names[i])]
  donor_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data_scaled[ , -1])))
donor_data_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
donor_data_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[ , which(sub("^([^_]*).*", "\\1", colnames(data_scaled)) %in% names[i])]
  donor_data_scaled[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data_scaled) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data_norm[ , -1])))
donor_data_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
donor_data_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[ , which(sub("^([^_]*).*", "\\1", colnames(data_norm)) %in% names[i])]
  donor_data_norm[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data_norm) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data_mode[ , -1])))
donor_data_mode <- data.frame(matrix(nrow = nrow(data_mode), ncol = length(names) + 1))
donor_data_mode$X1 <- data_mode$size
for (i in seq_along(names)) {
  data_sub <- data_mode[ , which(sub("^([^_]*).*", "\\1", colnames(data_mode)) %in% names[i])]
  donor_data_mode[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data_mode) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data[ , -1])))
fraction_data <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
fraction_data$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data)) %in% names[i])]
  fraction_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_scaled[ , -1])))
fraction_data_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
fraction_data_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_scaled)) %in% names[i])
  ]
  fraction_data_scaled[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data_scaled) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_norm[ , -1])))
fraction_data_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
fraction_data_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_norm)) %in% names[i])
  ]
  fraction_data_norm[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data_norm) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_mode[ , -1])))
fraction_data_mode <- data.frame(matrix(nrow = nrow(data_mode), ncol = length(names) + 1))
fraction_data_mode$X1 <- data_mode$size
for (i in seq_along(names)) {
  data_sub <- data_mode[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_mode)) %in% names[i])
  ]
  fraction_data_mode[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data_mode) <- c("size", names)
```

# Distribution Plots

We only plot sizes up to 500nm.
Data objects without any limit on size are also provided for browsing and download.

## All Data

### Unnormalised

```{r}
data_sub <- data[1:which(data$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, Concentration, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  theme(legend.position = "none") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 5e+10),
    limits = c(0, max(molten_data$Concentration)),
    labels = function(x) format(x, scientific = TRUE)
  )

datatable_download(data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- data_scaled[1:which(data_scaled$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, Concentration, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  theme(legend.position = "none") +
  xlab("Diameter (nm)") +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.005),
    limits = c(0, max(molten_data$Concentration))
  )

datatable_download(data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- data_norm[1:which(data_norm$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "relative_proportion", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, relative_proportion, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(
      ymin = relative_proportion - SD, ymax = relative_proportion + SD
    ),
    size = 0.2,
    position = position_dodge(0.05),
    color = "darkgrey"
  ) +
  theme_light() +
  theme(legend.position = "none") +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )

datatable_download(data_norm)
```

### Normalised by Mode

```{r}
data_sub <- data_mode[1:which(data_mode$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, Concentration, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  theme(legend.position = "none") +
  ylab("Normalised Relative Concentration") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.1),
    limits = c(0, max(molten_data$Concentration))
  )

datatable_download(data_mode)
```

## Donor by Fraction

### Unnormalised

```{r}
data_sub <- donor_fraction_data[1:which(donor_fraction_data$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, Concentration, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 5e+10),
    limits = c(0, max(molten_data$Concentration)),
    labels = function(x) format(x, scientific = TRUE)
  )

datatable_download(donor_fraction_data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- donor_fraction_data_scaled[1:which(donor_fraction_data_scaled$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, Concentration, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  xlab("Diameter (nm)") +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.005),
    limits = c(0, max(molten_data$Concentration))
  )

datatable_download(donor_fraction_data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- donor_fraction_data_norm[1:which(donor_fraction_data_norm$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "relative_proportion", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, relative_proportion, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(
      ymin = relative_proportion - SD, ymax = relative_proportion + SD
    ),
    size = 0.2,
    position = position_dodge(0.05),
    color = "darkgrey"
  ) +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )

datatable_download(donor_fraction_data_norm)
```

### Normalised by Mode

```{r}
data_sub <- donor_fraction_data_mode[1:which(donor_fraction_data_mode$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Sample)))

ggplot(molten_data, aes(Size, Concentration, group = Sample, color = Sample)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  ylab("Normalised Relative Concentration") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.1),
    limits = c(0, max(molten_data$Concentration))
  )

datatable_download(donor_fraction_data_mode)
```

## By Donor

```{r}
names(colour) <- colnames(donor_data[ , -1])
```

### Unnormalised

```{r}
data_sub <- donor_data[1:which(donor_data$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Donor", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Donor)))

ggplot(molten_data, aes(Size, Concentration, group = Donor, color = Donor)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 5e+10),
    limits = c(0, max(molten_data$Concentration)),
    labels = function(x) format(x, scientific = TRUE)
  ) +
  scale_colour_manual(values = colour)

datatable_download(donor_data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- donor_data_scaled[1:which(donor_data_scaled$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Donor", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Donor)))

ggplot(molten_data, aes(Size, Concentration, group = Donor, color = Donor)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  xlab("Diameter (nm)") +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.001),
    limits = c(0, max(molten_data$Concentration))
  )
  scale_colour_manual(values = colour)

datatable_download(donor_data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- donor_data_norm[1:which(donor_data_norm$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Donor", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "relative_proportion", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Donor)))

ggplot(molten_data, aes(Size, relative_proportion, group = Donor, color = Donor)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(
      ymin = relative_proportion - SD, ymax = relative_proportion + SD
    ),
    size = 0.2,
    position = position_dodge(0.05),
    color = "darkgrey"
  ) +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Donor (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  ) +
  scale_colour_manual(values = colour)

datatable_download(donor_data_norm)
```

### Normalised by Mode

```{r}
data_sub <- donor_data_mode[1:which(donor_data_mode$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Donor", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Donor)))

ggplot(molten_data, aes(Size, Concentration, group = Donor, color = Donor)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  ylab("Normalised Relative Concentration") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.1),
    limits = c(0, max(molten_data$Concentration))
  ) +
  scale_colour_manual(values = colour)

datatable_download(donor_data_mode)
```

## By Fraction

```{r}
names(colour) <- colnames(fraction_data[ , -1])
```

### Unnormalised

```{r}
data_sub <- fraction_data[1:which(fraction_data$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Fraction", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Fraction)))

ggplot(molten_data, aes(Size, Concentration, group = Fraction, color = Fraction)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 5e+10),
    limits = c(0, max(molten_data$Concentration)),
    labels = function(x) format(x, scientific = TRUE)
  ) +
  scale_colour_manual(values = colour)

datatable_download(fraction_data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- fraction_data_scaled[1:which(fraction_data_scaled$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Fraction", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Fraction)))

ggplot(molten_data, aes(Size, Concentration, group = Fraction, color = Fraction)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  xlab("Diameter (nm)") +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.002),
    limits = c(0, max(molten_data$Concentration))
  ) +
  scale_colour_manual(values = colour)

datatable_download(fraction_data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- fraction_data_norm[1:which(fraction_data_norm$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Fraction", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "relative_proportion", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Fraction)))

ggplot(molten_data, aes(Size, relative_proportion, group = Fraction, color = Fraction)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(
      ymin = relative_proportion - SD, ymax = relative_proportion + SD
    ),
    size = 0.2,
    position = position_dodge(0.05),
    color = "darkgrey"
  ) +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Fraction (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  ) +
  scale_colour_manual(values = colour)

datatable_download(fraction_data_norm)
```

### Normalised by Mode

```{r}
data_sub <- fraction_data_mode[1:which(fraction_data_mode$size == 597.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Fraction", "Concentration")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)

stats <- data_summary(molten_data, "Concentration", "Size")
molten_data$SD <- rep(stats$sd, length(unique(molten_data$Fraction)))

ggplot(molten_data, aes(Size, Concentration, group = Fraction, color = Fraction)) +
  geom_smooth(se = FALSE, span = 0.2) +
  geom_point(size = 0.75) +
  geom_errorbar(
    aes(ymin = Concentration - SD, ymax = Concentration + SD), size = 0.2,
    position = position_dodge(0.05), color = "darkgrey"
  ) +
  theme_light() +
  ylab("Normalised Relative Concentration") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Concentration), 0.1),
    limits = c(0, max(molten_data$Concentration))
  ) +
  scale_colour_manual(values = colour)

datatable_download(fraction_data_mode)
```

# Bar Plots

## All Data

### Unnormalised

```{r}
data_sub <- colSums(data[ , -1])
data_sub <- data.frame(Sample = names(data_sub), Concentration = data_sub)

ggplot(data_sub, aes(Sample, Concentration, fill = Sample)) +
  geom_bar(stat = "identity") +
  theme_light() +
  theme(legend.position = "none") +
  scale_y_continuous(
    breaks = seq(0, max(data_sub$Concentration), 2.5e+12),
    limits = c(0, max(data_sub$Concentration))
  ) +
  theme(axis.text.x = element_blank())
```

### Normalised By Tissue Amt.

```{r}
data_sub <- colSums(data_scaled[ , -1])
data_sub <- data.frame(Sample = names(data_sub), Concentration = data_sub)

ggplot(data_sub, aes(Sample, Concentration, fill = Sample)) +
  geom_bar(stat = "identity") +
  theme_light() +
  theme(legend.position = "none") +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_y_continuous(
    breaks = seq(0, max(data_sub$Concentration), 0.1),
    limits = c(0, max(data_sub$Concentration))
  ) +
  theme(axis.text.x = element_blank())
```

## Donor by Fraction

### Unnormalised

```{r}
data_sub <- colSums(donor_fraction_data[ , -1])
data_sub <- data.frame(Sample = names(data_sub), Concentration = data_sub)

ggplot(data_sub, aes(Sample, Concentration, fill = Sample)) +
  geom_bar(stat = "identity") +
  theme_light() +
  scale_y_continuous(
    breaks = seq(0, max(data_sub$Concentration), 2.5e+12),
    limits = c(0, max(data_sub$Concentration))
  ) +
  theme(axis.text.x = element_blank())
```

### Normalised By Tissue Amt.

```{r}
data_sub <- colSums(donor_fraction_data_scaled[ , -1])
data_sub <- data.frame(Sample = names(data_sub), Concentration = data_sub)

ggplot(data_sub, aes(Sample, Concentration, fill = Sample)) +
  geom_bar(stat = "identity") +
  theme_light() +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_y_continuous(
    breaks = seq(0, max(data_sub$Concentration), 0.1),
    limits = c(0, max(data_sub$Concentration))
  ) +
  theme(axis.text.x = element_blank())
```

## By Donor

```{r}
names(colour) <- colnames(donor_data[ , -1])
```

### Unnormalised

```{r}
data_sub <- colSums(donor_data[ , -1])
data_sub <- data.frame(Donor = names(data_sub), Concentration = data_sub)

ggplot(data_sub, aes(Donor, Concentration, fill = Donor)) +
  geom_bar(stat = "identity") +
  theme_light() +
  theme(legend.position = "none") +
  scale_y_continuous(
    breaks = seq(0, max(data_sub$Concentration) + 5e+11, 5e+11),
    limits = c(0, max(data_sub$Concentration))
  ) +
  scale_fill_manual(values = colour)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- colSums(donor_data_scaled[ , -1])
data_sub <- data.frame(Donor = names(data_sub), Concentration = data_sub)

ggplot(data_sub, aes(Donor, Concentration, fill = Donor)) +
  geom_bar(stat = "identity") +
  theme_light() +
  theme(legend.position = "none") +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_y_continuous(
    breaks = seq(0, max(data_sub$Concentration) + 0.05, 0.05),
    limits = c(0, max(data_sub$Concentration))
  ) +
  scale_fill_manual(values = colour)
```

## By Fraction

### Unnormalised

```{r}
data_sub <- colSums(donor_fraction_data[ , -1])
data_sub <- data.frame(Sample = names(data_sub), Concentration = data_sub)
data_sub$Donor <- sub("^([^_]*).*", "\\1", data_sub$Sample)
data_sub$Fraction <- sub("^[^_]*_([^_]*).*", "\\1", data_sub$Sample)

ggplot(data_sub, aes(Fraction, Concentration)) +
  geom_bar(stat = "summary", fill = "#F2F3F4", colour = "black") +
  geom_errorbar(aes(width = 0.75), stat = "summary") +
  geom_beeswarm(aes(fill = Donor), size = 2, pch = 21, cex = 2) +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(data_sub$Concentration) + 1e+12, 1e+12)) +
  scale_fill_manual(values = colour)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- colSums(donor_fraction_data_scaled[ , -1])
data_sub <- data.frame(Sample = names(data_sub), Concentration = data_sub)
data_sub$Donor <- sub("^([^_]*).*", "\\1", data_sub$Sample)
data_sub$Fraction <- sub("^[^_]*_([^_]*).*", "\\1", data_sub$Sample)

ggplot(data_sub, aes(Fraction, Concentration)) +
  geom_bar(stat = "summary", fill = "#F2F3F4", colour = "black") +
  geom_errorbar(aes(width = 0.75), stat = "summary") +
  geom_beeswarm(aes(fill = Donor), size = 2, pch = 21, cex = 2) +
  theme_light() +
  ylab("Concentration (billion particles / mg brain tissue)") +
  scale_y_continuous(breaks = seq(0, max(data_sub$Concentration) + 0.1, 0.1)) +
  scale_fill_manual(values = colour)
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
