---
title: "Zetaview Analysis"
author:
  - name: "Emir, Pal, and Steph"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-CBD-exosome-proteins.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "zetaview-analysis.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-CBD-exosome-proteins](https://github.com/eturkes/AD-CBD-exosome-proteins).*
*Please email for access.*

The data here will be referenced using the name `zetaview`.
It is particle count data from ZetaView of exosomes derived from post-mortem brains of 7 donors with Alzheimer's Disease (AD).
Exosomes were fractionated into 8 fractions (plus 10k pellet) for each donor.
Three technical replicates per fraction was run through the Zetaview.

```{r}
# Some standard boilerplate.
# --------------------------
# Copyright 2019-2022 Emir Turkes, Guar Pallavi, Stephanie Fowler, UK DRI at UCL, Columbia
# University Medical Center
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
packages <- c("conflicted", "DT", "ggplot2", "reshape2")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("utils.R"))

data_name <- "zetaview"
`%notin%` <- Negate(`%in%`)
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up data.

cache_dir <- file.path("..", "tmp", "cache", data_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# ----------------------------------------------------------
# --------------------------
```

# Prep

The data is spread out across many files, so we first aggregate them to obtain a single data object with counts per particle size.

```{r}
file_list <- list.files(file.path(assets_dir, data_name))

data <- data.frame(matrix(nrow = 1200, ncol = length(file_list)) + 1)

for (i in seq_along(file_list)) {
  file <- read.delim(
    file.path(assets_dir, data_name, file_list[i]), skip = 70
  )
  if (i == 1) {
    data[ , i] <- file[1:1200, 1]
  }
  data[ , i + 1] <- file[1:1200, 2]
}

colnames(data) <- c(
  "size",
    paste(
      sub("^[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sub("^[^_]*_[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sub("^[^_]*_[^_]*_[^_]*_[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sep = "_"
  )
)

datatable_download(data)
```

# Distribution Plots

```{r}
non_zero <- which(rowSums(data[ , -1]) > 0)
data <- data[min(non_zero):max(non_zero), ]

molten_data <- melt(data[ , -1], id.vars = NULL)
colnames(molten_data) <- c("sample", "frequency")
molten_data$size <- rep(data$size, ncol(data) - 1)

donor_fraction <- unique(sub("_[^_]+$", "", colnames(data[ , -1])))
aggregated_data <- data.frame(matrix(nrow = nrow(data), ncol = length(donor_fraction) + 1))
aggregated_data$X1 <- data$size
for (i in seq_along(donor_fraction)) {
  data_sub <- data[ , which(sub("_[^_]+$", "", colnames(data)) %in% donor_fraction[i])]
  aggregated_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(aggregated_data) <- c("size", donor_fraction)

aggregated_molten_data <- melt(aggregated_data[ , -1], id.vars = NULL)
colnames(aggregated_molten_data) <- c("sample", "frequency")
aggregated_molten_data$size <- rep(aggregated_data$size, ncol(aggregated_data) - 1)
```

## Smoothed

```{r}
ggplot(molten_data, aes(size, frequency, color = sample)) +
  geom_smooth(span = 0.1) +
  theme_light() +
  theme(legend.position = "none")
```

## Without Smoothing

```{r}
ggplot(molten_data, aes(size, frequency, color = sample)) +
  geom_line() +
  theme_light() +
  theme(legend.position = "none")
```

## Aggregated Smoothed

```{r}
ggplot(aggregated_molten_data, aes(size, frequency, color = sample)) +
  geom_smooth(span = 0.1) +
  theme_light()
```

## Aggregated Without Smoothing

```{r}
ggplot(aggregated_molten_data, aes(size, frequency, color = sample)) +
  geom_line() +
  theme_light()
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
