---
title: "Zetaview Analysis"
author:
  - name: "Emir, Pal, and Steph"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: "../AD-CBD-exosome-proteins.bib"
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "results", "zetaview-analysis.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [AD-CBD-exosome-proteins](https://github.com/eturkes/AD-CBD-exosome-proteins).*
*Please email for access.*

The data here will be referenced using the name `zetaview`.
It is particle count data from ZetaView of exosomes derived from post-mortem brains of 7 donors with Alzheimer's Disease (AD).
Exosomes were fractionated into 8 fractions (plus 10k pellet, which are excluded from downstream analysis) for each donor.
Three technical replicates per fraction was run through the Zetaview.

```{r}
# Some standard boilerplate.
# --------------------------
# Copyright 2019-2022 Emir Turkes, Guar Pallavi, Stephanie Fowler, UK DRI at UCL, Columbia
# University Medical Center
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
packages <- c("conflicted", "DT", "ggplot2", "reshape2", "scales")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("utils.R"))

data_name <- "zetaview"
`%notin%` <- Negate(`%in%`)
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up data.

cache_dir <- file.path("..", "tmp", "cache", data_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# ----------------------------------------------------------
# --------------------------
```

# Prep

The data is spread out across many files, so we first aggregate them to obtain a single data object with counts per particle size.
We also create variations of this object that are aggregated by the mean of values by donor and fraction as well as various normalisation schemes found in following sections.

```{r}
file_list <- list.files(file.path(assets_dir, data_name, "raw-data"))
file_list <- file_list[-grep("10K", file_list)] # Remove 10K pellets.

data <- data.frame(matrix(nrow = 1200, ncol = length(file_list)) + 1)

for (i in seq_along(file_list)) {
  file <- read.delim(
    file.path(assets_dir, data_name, "raw-data", file_list[i]), skip = 70
  )
  if (i == 1) {
    data[ , i] <- file[1:1200, 1]
  }
  data[ , i + 1] <- file[1:1200, 2]
}

colnames(data) <- c(
  "size",
    paste(
      sub("^[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sub("^[^_]*_[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sub("^[^_]*_[^_]*_[^_]*_[^_]*_[^_]*_([^_]*).*", "\\1", file_list),
      sep = "_"
  )
)

non_zero <- which(rowSums(data[ , -1]) > 0)
data <- data[min(non_zero):max(non_zero), ]

# Scaling factor based on tissue amount used.
# -------------------------------------------
scaling_factor <- data.frame(amount = c(1.29, 1.21, 1.19, 1.05, 1.01, 1.15, 1.2))
scaling_factor <- min(scaling_factor, na.rm = TRUE) / scaling_factor
rownames(scaling_factor) <- unique(sub("^([^_]*).*", "\\1", colnames(data[ , -1])))
# -------------------------------------------

data_scaled <- data
for (i in seq_along(rownames(scaling_factor))) {
  for (j in seq_along(colnames(data_scaled))) {
    if (sub("^([^_]*).*", "\\1", colnames(data))[j] == rownames(scaling_factor)[i])
      data_scaled[j] <- data_scaled[j] * scaling_factor[i, ]
  }
}

data_norm <- data
for (i in seq_along(colnames(data_norm))) {
  if (i != 1) {
    data_norm[i] <- (data_norm[i] / sum(data_norm[i])) * 100
  }
}

names <- unique(sub("_[^_]+$", "", colnames(data[ , -1])))
donor_fraction_data <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
donor_fraction_data$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("_[^_]+$", "", colnames(data)) %in% names[i])]
  donor_fraction_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data) <- c("size", names)

names <- unique(sub("_[^_]+$", "", colnames(data_scaled[ , -1])))
donor_fraction_data_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
donor_fraction_data_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[ , which(sub("_[^_]+$", "", colnames(data_scaled)) %in% names[i])]
  donor_fraction_data_scaled[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data_scaled) <- c("size", names)

names <- unique(sub("_[^_]+$", "", colnames(data_norm[ , -1])))
donor_fraction_data_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
donor_fraction_data_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[ , which(sub("_[^_]+$", "", colnames(data_norm)) %in% names[i])]
  donor_fraction_data_norm[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_fraction_data_norm) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data[ , -1])))
donor_data <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
donor_data$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("^([^_]*).*", "\\1", colnames(data)) %in% names[i])]
  donor_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data_scaled[ , -1])))
donor_data_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
donor_data_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[ , which(sub("^([^_]*).*", "\\1", colnames(data_scaled)) %in% names[i])]
  donor_data_scaled[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data_scaled) <- c("size", names)

names <- unique(sub("^([^_]*).*", "\\1", colnames(data_norm[ , -1])))
donor_data_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
donor_data_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[ , which(sub("^([^_]*).*", "\\1", colnames(data_norm)) %in% names[i])]
  donor_data_norm[ , i + 1] <- rowMeans(data_sub)
}
colnames(donor_data_norm) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data[ , -1])))
fraction_data <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
fraction_data$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data)) %in% names[i])]
  fraction_data[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_scaled[ , -1])))
fraction_data_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
fraction_data_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_scaled)) %in% names[i])
  ]
  fraction_data_scaled[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data_scaled) <- c("size", names)

names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_norm[ , -1])))
fraction_data_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
fraction_data_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_norm)) %in% names[i])
  ]
  fraction_data_norm[ , i + 1] <- rowMeans(data_sub)
}
colnames(fraction_data_norm) <- c("size", names)
```

# Distribution Plots

We only plot sizes up to 300nm as counts quickly approach zero from there.
Data objects without any limit on size are also provided for browsing and download.

## All Data

### Unnormalised

```{r}
data_sub <- data[1:which(data$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth() +
  theme_light() +
  theme(legend.position = "none") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  theme(legend.position = "none") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- data_scaled[1:which(data_scaled$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth() +
  theme_light() +
  theme(legend.position = "none") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  theme(legend.position = "none") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- data_norm[1:which(data_norm$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_smooth() +
  theme_light() +
  theme(legend.position = "none") +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_line() +
  theme_light() +
  theme(legend.position = "none") +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Data Table**

```{r}
datatable_download(data_norm)
```

## Donor by Fraction

### Unnormalised

```{r}
data_sub <- donor_fraction_data[1:which(donor_fraction_data$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(donor_fraction_data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- donor_fraction_data_scaled[1:which(donor_fraction_data_scaled$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(donor_fraction_data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- donor_fraction_data_norm[1:which(donor_fraction_data_norm$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_smooth() +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_line() +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Data Table**

```{r}
datatable_download(donor_fraction_data_norm)
```

## By Donor

### Unnormalised

```{r}
data_sub <- donor_data[1:which(donor_data$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(donor_data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- donor_data_scaled[1:which(donor_data_scaled$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(donor_data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- donor_data_norm[1:which(donor_data_norm$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_smooth() +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_line() +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Data Table**

```{r}
datatable_download(donor_data_norm)
```

## By Fraction

### Unnormalised

```{r}
data_sub <- fraction_data[1:which(fraction_data$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth(span = 0.1) +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(fraction_data)
```

### Normalised By Tissue Amt.

```{r}
data_sub <- fraction_data_scaled[1:which(fraction_data_scaled$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "Frequency")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_smooth(span = 0.1) +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
  geom_line() +
  theme_light() +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
  )
```

**Data Table**

```{r}
datatable_download(fraction_data_scaled)
```

### Proportion of Total Counts

```{r}
data_sub <- fraction_data_norm[1:which(fraction_data_norm$size == 297.5), ]
molten_data <- melt(data_sub[ , -1], id.vars = NULL)
colnames(molten_data) <- c("Sample", "relative_proportion")
molten_data$Size <- rep(data_sub$size, ncol(data_sub) - 1)
```

**Smoothed**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_smooth(span = 0.1) +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Without Smoothing**

```{r}
ggplot(molten_data, aes(Size, relative_proportion, color = Sample)) +
  geom_line() +
  theme_light() +
  ylab("Relative Proportion of Total Counts Per Sample (%)") +
  xlab("Diameter (nm)") +
  scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
  scale_y_continuous(
    breaks = seq(0, max(molten_data$relative_proportion), 0.5),
    limits = c(0, max(molten_data$relative_proportion))
  )
```

**Data Table**

```{r}
datatable_download(fraction_data_norm)
```

# Summary Stats

Here we calculate the mean, median, mode and percentiles of each distribution.
We only focus on the "By Fraction" data as it is most relevant.

## By Fraction

### Unnormalised

```{r}
names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data[ , -1])))
fraction_sums <- data.frame(matrix(nrow = nrow(data), ncol = length(names) + 1))
fraction_sums$X1 <- data$size
for (i in seq_along(names)) {
  data_sub <- data[ , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data)) %in% names[i])]
  fraction_sums[ , i + 1] <- rowSums(data_sub)
}
colnames(fraction_sums) <- c("size", names)

stats <- data.frame(matrix(nrow = 6, ncol = ncol(fraction_sums)- 1))
colnames(stats) <- colnames(fraction_sums[ , -1])
rownames(stats) <- c("Mean", "Median", "Mode", "D10", "D50", "D90")

counts_per_size <- vector("list", ncol(fraction_sums)- 1)
for (i in seq_along(counts_per_size)) {
  counts_per_size[[i]] <- rep(fraction_sums[ , 1], fraction_sums[ , i + 1])
  stats[1, i] <- mean(counts_per_size[[i]])
  stats[2, i] <- median(counts_per_size[[i]])
  stats[3, i] <- as.numeric(names(table(counts_per_size[[i]]))[
    table(counts_per_size[[i]]) == max(table(counts_per_size[[i]]))
  ])[1]
  stats[4:6, i] <- as.numeric(quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9)))
}

datatable_download(stats)

molten_data <- as.data.frame(t(stats))
molten_data$Fraction <- rownames(molten_data)
molten_data <- melt(molten_data, id.vars = "Fraction")
colnames(molten_data) <- c("Fraction", "Statistic", "Diameter")

ggplot(molten_data, aes(Fraction, Diameter, fill = Statistic)) +
  geom_bar(stat = "identity", colour = "black", position = "dodge", width = 0.75) +
  ylab("Diameter (nm)") +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(molten_data$Diameter), 10))

stats_scaled <- stats
for (i in seq(nrow(stats))) {
  stats_scaled[i, ] <- stats[i, ] - min(stats[i, ])
  stats_scaled[i, ] <- stats_scaled[i, ] / max(stats_scaled[i, ])
}

molten_data <- as.data.frame(t(stats_scaled))
molten_data$Fraction <- rownames(molten_data)
molten_data <- melt(molten_data, id.vars = "Fraction")
colnames(molten_data) <- c("Fraction", "Statistic", "Diameter")

ggplot(molten_data, aes(Fraction, Diameter, fill = Statistic)) +
  geom_bar(stat = "identity", colour = "black", position = "dodge", width = 0.75) +
  ylab("Relative Difference in Diameter") +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(molten_data$Diameter), 0.05))

for (i in seq(ncol(stats))) {
  data_sub1 <- fraction_data[1:which(fraction_data$size == stats[4, i]), ]
  data_sub1[nrow(data_sub1) , 1] <- stats[4, i] - 10e-6
  data_sub2 <- fraction_data[
    which(fraction_data$size == stats[4, i]):which(fraction_data$size == stats[5, i]),
  ]
  data_sub2[nrow(data_sub2) , 1] <- stats[5, i] - 10e-6
  data_sub3 <- fraction_data[
    which(fraction_data$size == stats[5, i]):which(fraction_data$size == stats[6, i]),
  ]
  data_sub3[nrow(data_sub3) , 1] <- stats[6, i] - 10e-6
  data_sub4 <- fraction_data[
    which(fraction_data$size == stats[6, i]):which(fraction_data_norm$size == 297.5),
  ]
  data_sub <- do.call("rbind", list(data_sub1, data_sub2, data_sub3, data_sub4))

  molten_data <- melt(data_sub[ , -1], id.vars = NULL)
  molten_data <- molten_data[molten_data$variable == colnames(stats)[i], ]
  colnames(molten_data) <- c("Sample", "Frequency")
  molten_data$Size <- data_sub$size
  molten_data$percentiles <- factor(
    findInterval(molten_data$Size, quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9)))
  )

  plot(
    ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
      geom_area(aes(fill = percentiles), outline.type = "full") +
      scale_x_continuous(
        sec.axis = dup_axis(
          name = "Percentile", breaks = quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9))
        )
      ) +
      scale_fill_brewer(guide = "none") +
      scale_color_manual(values = hue_pal()(8)[i]) +
      theme_light() +
      geom_vline(xintercept = stats[4, i]) +
      geom_vline(xintercept = stats[5, i]) +
      geom_vline(xintercept = stats[6, i]) +
      guides(color = guide_legend(override.aes = list(fill = NA))) +
      xlab("Diameter (nm)") +
      scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
      scale_y_continuous(
        breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
      )
  )
}
```

### Normalised By Tissue Amt.

```{r}
names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_scaled[ , -1])))
fraction_sums_scaled <- data.frame(matrix(nrow = nrow(data_scaled), ncol = length(names) + 1))
fraction_sums_scaled$X1 <- data_scaled$size
for (i in seq_along(names)) {
  data_sub <- data_scaled[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_scaled)) %in% names[i])
  ]
  fraction_sums_scaled[ , i + 1] <- rowSums(data_sub)
}
colnames(fraction_sums_scaled) <- c("size", names)

stats <- data.frame(matrix(nrow = 6, ncol = ncol(fraction_sums_scaled)- 1))
colnames(stats) <- colnames(fraction_sums_scaled[ , -1])
rownames(stats) <- c("Mean", "Median", "Mode", "D10", "D50", "D90")

counts_per_size <- vector("list", ncol(fraction_sums_scaled)- 1)
for (i in seq_along(counts_per_size)) {
  counts_per_size[[i]] <- rep(fraction_sums_scaled[ , 1], fraction_sums_scaled[ , i + 1])
  stats[1, i] <- mean(counts_per_size[[i]])
  stats[2, i] <- median(counts_per_size[[i]])
  stats[3, i] <- as.numeric(names(table(counts_per_size[[i]]))[
    table(counts_per_size[[i]]) == max(table(counts_per_size[[i]]))
  ])[1]
  stats[4:6, i] <- as.numeric(quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9)))
}

datatable_download(stats)

molten_data <- as.data.frame(t(stats))
molten_data$Fraction <- rownames(molten_data)
molten_data <- melt(molten_data, id.vars = "Fraction")
colnames(molten_data) <- c("Fraction", "Statistic", "Diameter")

ggplot(molten_data, aes(Fraction, Diameter, fill = Statistic)) +
  geom_bar(stat = "identity", colour = "black", position = "dodge", width = 0.75) +
  ylab("Diameter (nm)") +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(molten_data$Diameter), 10))

stats_scaled <- stats
for (i in seq(nrow(stats))) {
  stats_scaled[i, ] <- stats[i, ] - min(stats[i, ])
  stats_scaled[i, ] <- stats_scaled[i, ] / max(stats_scaled[i, ])
}

molten_data <- as.data.frame(t(stats_scaled))
molten_data$Fraction <- rownames(molten_data)
molten_data <- melt(molten_data, id.vars = "Fraction")
colnames(molten_data) <- c("Fraction", "Statistic", "Diameter")

ggplot(molten_data, aes(Fraction, Diameter, fill = Statistic)) +
  geom_bar(stat = "identity", colour = "black", position = "dodge", width = 0.75) +
  ylab("Relative Difference in Diameter") +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(molten_data$Diameter), 0.05))

for (i in seq(ncol(stats))) {
  data_sub1 <- fraction_data_scaled[1:which(fraction_data_scaled$size == stats[4, i]), ]
  data_sub1[nrow(data_sub1) , 1] <- stats[4, i] - 10e-6
  data_sub2 <- fraction_data_scaled[
    which(fraction_data_scaled$size == stats[4, i]):which(fraction_data_scaled$size == stats[5, i]),
  ]
  data_sub2[nrow(data_sub2) , 1] <- stats[5, i] - 10e-6
  data_sub3 <- fraction_data_scaled[
    which(fraction_data_scaled$size == stats[5, i]):which(fraction_data_scaled$size == stats[6, i]),
  ]
  data_sub3[nrow(data_sub3) , 1] <- stats[6, i] - 10e-6
  data_sub4 <- fraction_data_scaled[
    which(fraction_data_scaled$size == stats[6, i]):which(fraction_data_scaled$size == 297.5),
  ]
  data_sub <- do.call("rbind", list(data_sub1, data_sub2, data_sub3, data_sub4))

  molten_data <- melt(data_sub[ , -1], id.vars = NULL)
  molten_data <- molten_data[molten_data$variable == colnames(stats)[i], ]
  colnames(molten_data) <- c("Sample", "Frequency")
  molten_data$Size <- data_sub$size
  molten_data$percentiles <- factor(
    findInterval(molten_data$Size, quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9)))
  )

  plot(
    ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
      geom_area(aes(fill = percentiles), outline.type = "full") +
      scale_x_continuous(
        sec.axis = dup_axis(
          name = "Percentile", breaks = quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9))
        )
      ) +
      scale_fill_brewer(guide = "none") +
      scale_color_manual(values = hue_pal()(8)[i]) +
      theme_light() +
      geom_vline(xintercept = stats[4, i]) +
      geom_vline(xintercept = stats[5, i]) +
      geom_vline(xintercept = stats[6, i]) +
      guides(color = guide_legend(override.aes = list(fill = NA))) +
      xlab("Diameter (nm)") +
      scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
      scale_y_continuous(
        breaks = seq(0, max(molten_data$Frequency), 5), limits = c(0, max(molten_data$Frequency))
      )
  )
}
```

### Proportion of Total Counts

```{r}
names <- unique(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_norm[ , -1])))
fraction_sums_norm <- data.frame(matrix(nrow = nrow(data_norm), ncol = length(names) + 1))
fraction_sums_norm$X1 <- data_norm$size
for (i in seq_along(names)) {
  data_sub <- data_norm[
    , which(sub("^[^_]*_([^_]*).*", "\\1", colnames(data_norm)) %in% names[i])
  ]
  fraction_sums_norm[ , i + 1] <- rowSums(data_sub)
}
colnames(fraction_sums_norm) <- c("size", names)

stats <- data.frame(matrix(nrow = 6, ncol = ncol(fraction_sums_norm)- 1))
colnames(stats) <- colnames(fraction_sums_norm[ , -1])
rownames(stats) <- c("Mean", "Median", "Mode", "D10", "D50", "D90")

counts_per_size <- vector("list", ncol(fraction_sums_norm)- 1)
for (i in seq_along(counts_per_size)) {
  counts_per_size[[i]] <- rep(fraction_sums_norm[ , 1], fraction_sums_norm[ , i + 1])
  stats[1, i] <- mean(counts_per_size[[i]])
  stats[2, i] <- median(counts_per_size[[i]])
  stats[3, i] <- as.numeric(names(table(counts_per_size[[i]]))[
    table(counts_per_size[[i]]) == max(table(counts_per_size[[i]]))
  ])[1]
  stats[4:6, i] <- as.numeric(quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9)))
}

datatable_download(stats)

molten_data <- as.data.frame(t(stats))
molten_data$Fraction <- rownames(molten_data)
molten_data <- melt(molten_data, id.vars = "Fraction")
colnames(molten_data) <- c("Fraction", "Statistic", "Diameter")

ggplot(molten_data, aes(Fraction, Diameter, fill = Statistic)) +
  geom_bar(stat = "identity", colour = "black", position = "dodge", width = 0.75) +
  ylab("Diameter (nm)") +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(molten_data$Diameter), 10))

stats_scaled <- stats
for (i in seq(nrow(stats))) {
  stats_scaled[i, ] <- stats[i, ] - min(stats[i, ])
  stats_scaled[i, ] <- stats_scaled[i, ] / max(stats_scaled[i, ])
}

molten_data <- as.data.frame(t(stats_scaled))
molten_data$Fraction <- rownames(molten_data)
molten_data <- melt(molten_data, id.vars = "Fraction")
colnames(molten_data) <- c("Fraction", "Statistic", "Diameter")

ggplot(molten_data, aes(Fraction, Diameter, fill = Statistic)) +
  geom_bar(stat = "identity", colour = "black", position = "dodge", width = 0.75) +
  ylab("Relative Difference in Diameter") +
  theme_light() +
  scale_y_continuous(breaks = seq(0, max(molten_data$Diameter), 0.05))

for (i in seq(ncol(stats))) {
  data_sub1 <- fraction_data_norm[1:which(fraction_data_norm$size == stats[4, i]), ]
  data_sub1[nrow(data_sub1) , 1] <- stats[4, i] - 10e-6
  data_sub2 <- fraction_data_norm[
    which(fraction_data_norm$size == stats[4, i]):which(fraction_data_norm$size == stats[5, i]),
  ]
  data_sub2[nrow(data_sub2) , 1] <- stats[5, i] - 10e-6
  data_sub3 <- fraction_data_norm[
    which(fraction_data_norm$size == stats[5, i]):which(fraction_data_norm$size == stats[6, i]),
  ]
  data_sub3[nrow(data_sub3) , 1] <- stats[6, i] - 10e-6
  data_sub4 <- fraction_data_norm[
    which(fraction_data_norm$size == stats[6, i]):which(fraction_data_norm$size == 297.5),
  ]
  data_sub <- do.call("rbind", list(data_sub1, data_sub2, data_sub3, data_sub4))

  molten_data <- melt(data_sub[ , -1], id.vars = NULL)
  molten_data <- molten_data[molten_data$variable == colnames(stats)[i], ]
  colnames(molten_data) <- c("Sample", "Frequency")
  molten_data$Size <- data_sub$size
  molten_data$percentiles <- factor(
    findInterval(molten_data$Size, quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9)))
  )

  plot(
    ggplot(molten_data, aes(Size, Frequency, color = Sample)) +
      geom_area(aes(fill = percentiles), outline.type = "full") +
      scale_x_continuous(
        sec.axis = dup_axis(
          name = "Percentile", breaks = quantile(counts_per_size[[i]], probs = c(0.1, 0.5, 0.9))
        )
      ) +
      scale_fill_brewer(guide = "none") +
      scale_color_manual(values = hue_pal()(8)[i]) +
      theme_light() +
      geom_vline(xintercept = stats[4, i]) +
      geom_vline(xintercept = stats[5, i]) +
      geom_vline(xintercept = stats[6, i]) +
      guides(color = guide_legend(override.aes = list(fill = NA))) +
      xlab("Diameter (nm)") +
      scale_x_continuous(breaks = seq(0, max(molten_data$Size), 20)) +
      scale_y_continuous(
        breaks = seq(0, max(molten_data$Frequency), 0.5), limits = c(0, max(molten_data$Frequency))
      ) +
      ylab("Relative Proportion of Total Counts Per Sample (%)")
  )
}
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
